---
title: "DATA 621 Homework 4"
author: "Critical Thinking Group 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '2'
    smart: no
  word_document:
    toc: yes
    toc_depth: '2'
---
\pagebreak
\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Prof. Dr. Nasrin Khansari\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
DATA 621 – Business Analytics and Data Mining\\
\medskip
Home Work 3\\
\medskip
Prepared by:\\
\medskip
Critical Thinking Group  1\\ 
\medskip
Vic Chan\\ 
\smallskip
Gehad Gad\\
\smallskip
Evan McLaughlin\\  
\smallskip
Bruno de Melo\\
\smallskip
Anjal Hussan\\
\smallskip
Zhouxin Shi\\
\smallskip
Sie Siong Wong\\
\end{center}

\pagebreak


```{R message=FALSE, warning=FALSE, echo=FALSE}
if (!require('ggplot2')) (install.packages('ggplot2'))
if(!require('dplyr')) (install.packages('dplyr'))
if(!require('tidyr')) (install.packages('tidyr'))
if(!require('purrr')) (install.packages('purrr'))

```


\newpage

# Overview

In this homework assignment, you will explore, analyze and model a data set containing approximately 8000 records representing a customer at an auto insurance company. Each record has two response variables. The first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.


# Statement of the Problem

The purpose of this report is to develop a binary logistic regression model to determine if the neighborhood will be at risk for high crime level.

The purpose of this report is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. 

# Data Exploration  

```{r echo=FALSE}

Train_Data <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW4/insurance_training_data.csv")

```


Let's take a look to the structure of our train data set


```{r}
str(Train_Data)
```
We can see that the training data has 8161 observations(rows) and  26 variables(columns).

We can notice that many features that are factors are imported as characters or doubles. Also, there may be some ordinal levels within some of the factors. 


```{r}
#Summary of each feature

summary(Train_Data) 
```

We can observe the followings:

**KIDSDRIV**: Max is 4

**AGE**: age is 16 is the youngest and oldest 81. There are 6 NA values

**HOMEKIDS**: Max is 5

**TRAVTIME**: 75% of the population is below 44 but the Max value is 142. It looks like there may be some outliers here. 

**TIF**: The majority of people are not long time customers

**CLM_FREQ**: Maximum is over 5 years

**MVR_PTS**: 75% have 3 or less, maximum is 13

**CAR_AGE**: Strange!. The minimum -3 and Max is 28. There are 510 NA values.


##Missing Values

```{r}
#missing values by column
colSums(is.na(Train_Data)) 
```

There are missing values in AGE (6), YOJ(454), and CAR_AGE(510)..


```{r}
Train_Data1<-select_if(Train_Data, is.numeric)
Train_Data1 %>%
  keep(is.numeric) %>%                     
  gather() %>%                            
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +  
    geom_density()  
```





