---
title: "DATA 621 Homework 4"
author: "Critical Thinking Group 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '2'
    smart: no
  word_document:
    toc: yes
    toc_depth: '2'
---
\pagebreak
\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Prof. Dr. Nasrin Khansari\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
DATA 621 – Business Analytics and Data Mining\\
\medskip
Home Work 4\\
\medskip
Prepared by:\\
\medskip
Critical Thinking Group  1\\ 
\medskip
Vic Chan\\ 
\smallskip
Gehad Gad\\
\smallskip
Evan McLaughlin\\  
\smallskip
Bruno de Melo\\
\smallskip
Anjal Hussan\\
\smallskip
Zhouxin Shi\\
\smallskip
Sie Siong Wong\\
\end{center}

\pagebreak


```{r message=FALSE, warning=FALSE, echo=FALSE}
if (!require('ggplot2')) (install.packages('ggplot2'))
if(!require('dplyr')) (install.packages('dplyr'))
if(!require('tidyverse')) (install.packages('tidyverse'))
if(!require('purrr')) (install.packages('purrr'))
if(!require('mice')) (install.packages('mice'))
if(!require('DataExplorer')) (install.packages('DataExplorer'))
if(!require('MASS')) (install.packages('MASS'))
if(!require('caret')) (install.packages('caret'))
if(!require('stats')) (install.packages('stats'))
if(!require('randomForest')) (install.packages('randomForest'))
if(!require('pROC')) (install.packages('pROC'))
if(!require('kableExtra')) (install.packages('kableExtra'))
if(!require('gridExtra')) (install.packages('gridExtra'))
if(!require('ggcorrplot')) (install.packages("ggcorrplot"))
if(!require('fitdistrplus')) (install.packages("fitdistrplus"))
```


\newpage

# Introduction

In this homework assignment, we will explore, analyze and model a data set containing approximately 8000 records representing a customer at an auto insurance company. Each record has two response variables. The first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.


# Statement of the Problem

The purpose of this report is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. 

# Data Exploration  

```{r echo=FALSE}
Train_Data <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW4/insurance_training_data.csv")
Eval_Data <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW4/insurance-evaluation-data.csv")
```


Let's take a look in the structure of our train data set - excluding the first column **index** which it is not to be used. Evaluation data set structure is similar to the train data set and will go through same


```{r echo=FALSE}
Train_Data <- Train_Data[,-1]
str(Train_Data)
Eval_Data <- Eval_Data[,-1]
```
We can see that the training data has 8161 observations(rows) and 25 variables (columns). Of these 25 columns, many are of factors type but were imported as characters or doubles - there will be properly converted in the preparation section. Also, there may be some ordinal levels within some of the factors.

Below we display a summary of each feature.


```{r echo=FALSE}
#Summary of each feature
summary(Train_Data) 
```

We can observe the followings:

**KIDSDRIV**: Max is 4

**AGE**: age is 16 is the youngest and oldest 81. There are 6 NA values

**HOMEKIDS**: Max is 5

**TRAVTIME**: 75% of the population is below 44 but the Max value is 142. It looks like there may be some outliers here.

**TIF**: The majority of people are not long time customers

**CLM_FREQ**: Maximum is over 5 years

**MVR_PTS**: 75% have 3 or less, maximum is 13

**CAR_AGE**: Strange!. The minimum -3 and Max is 28. There are 510 NA values. These negative values will have to be excluded from the analysis.

**INCOME** - **BLUEBOOK** - **HOME_VAL** - **OLDCLAIM** : These are numerical variables that need to be converted accordingly.

## Convertion to numerical

As can be seen below, these four features are now corrected represented.

```{r echo=FALSE}
Train_Data$INCOME <- gsub(",","",(Train_Data$INCOME))
Train_Data$INCOME <- sub('.', '', Train_Data$INCOME)
Train_Data$INCOME <-trimws(Train_Data$INCOME, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$INCOME <- as.numeric(Train_Data$INCOME)
Train_Data$HOME_VAL <- gsub(",","",(Train_Data$HOME_VAL))
Train_Data$HOME_VAL <- sub('.', '', Train_Data$HOME_VAL)
Train_Data$HOME_VAL <-trimws(Train_Data$HOME_VAL, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$HOME_VAL <- as.numeric(Train_Data$HOME_VAL)
#
Train_Data$BLUEBOOK <- gsub(",","",(Train_Data$BLUEBOOK))
Train_Data$BLUEBOOK <- sub('.', '', Train_Data$BLUEBOOK)
Train_Data$BLUEBOOK <-trimws(Train_Data$BLUEBOOK, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$BLUEBOOK <- as.numeric(Train_Data$BLUEBOOK)
#
Train_Data$OLDCLAIM <- gsub(",","",(Train_Data$OLDCLAIM))
Train_Data$OLDCLAIM <- sub('.', '', Train_Data$OLDCLAIM)
Train_Data$OLDCLAIM <-trimws(Train_Data$OLDCLAIM, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$OLDCLAIM <- as.numeric(Train_Data$OLDCLAIM)
#
Eval_Data$INCOME <- gsub(",","",(Eval_Data$INCOME))
Eval_Data$INCOME <- sub('.', '', Eval_Data$INCOME)
Eval_Data$INCOME <-trimws(Eval_Data$INCOME, which = c("both"), whitespace = "[ \t\r\n]")
Eval_Data$INCOME <- as.numeric(Eval_Data$INCOME)
Eval_Data$HOME_VAL <- gsub(",","",(Eval_Data$HOME_VAL))
Eval_Data$HOME_VAL <- sub('.', '', Eval_Data$HOME_VAL)
Eval_Data$HOME_VAL <-trimws(Eval_Data$HOME_VAL, which = c("both"), whitespace = "[ \t\r\n]")
Eval_Data$HOME_VAL <- as.numeric(Eval_Data$HOME_VAL)
#
Eval_Data$BLUEBOOK <- gsub(",","",(Eval_Data$BLUEBOOK))
Eval_Data$BLUEBOOK <- sub('.', '', Eval_Data$BLUEBOOK)
Eval_Data$BLUEBOOK <-trimws(Eval_Data$BLUEBOOK, which = c("both"), whitespace = "[ \t\r\n]")
Eval_Data$BLUEBOOK <- as.numeric(Eval_Data$BLUEBOOK)
#
Eval_Data$OLDCLAIM <- gsub(",","",(Eval_Data$OLDCLAIM))
Eval_Data$OLDCLAIM <- sub('.', '', Eval_Data$OLDCLAIM)
Eval_Data$OLDCLAIM <-trimws(Eval_Data$OLDCLAIM, which = c("both"), whitespace = "[ \t\r\n]")
Eval_Data$OLDCLAIM <- as.numeric(Eval_Data$OLDCLAIM)
#
summary(Train_Data[,c(7,9,16,20)])
```

## Missing Values

```{r echo=FALSE}
#missing values by column
colSums(is.na(Train_Data)) 
```

There are missing values in several variables for a total of 1,879 NA's or about 1% of the total dataset.

## Univariate Distribution - Histograms

Below the numeric feature distributions are displayed.

```{r message=FALSE, warning=FALSE, echo=FALSE}
Train_Data1<-dplyr::select_if(Train_Data, is.numeric)
Train_Data1 %>%
  keep(is.numeric) %>%                     
  tidyr::gather() %>%                            
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +  
    geom_density()  
```
We can see that AGE, BLUEBOOK, CAR_AGE, HOME_VAL, INCOME, TRAVTIME and YOJ resemblance somewhat a normal distribution while CLM_FREQ, HOMEKIDS, KIDSDRIV, MVR_PTS, OLDCLAIM, TARGET_AMT, TIF resemblance either a binomial or Poisson distribution.

Let's investigate using a qq_plot:

```{r message=FALSE, warning=FALSE, echo=FALSE}
Train_Data1<-select_if(Train_Data, is.numeric)
Train_Data1 %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(sample=value)) +                    
    facet_wrap(~ key, scales = "free") +  
      stat_qq(distribution = stats::qnorm) +
      stat_qq_line(distribution = stats::qnorm)
```

In order to descriptive the distribution, we used function 'descdist' from package 'fitdistrplus'. We display one output for illustrative purposes - on featue **KIDSDRIV**, and results for all other features are shown below:

```{r message=FALSE, warning=FALSE, echo=FALSE}
  descdist(Train_Data1[,3], discrete = TRUE)
```

We can observe the followings:

**AGE**: normal distribution

*BLUEBOOK**: quasi-normal/lognormal - skewed distribution with heavy tails

**CAR_AGE**: quasi-normal/lognormal - skewed distribution with high frequency of <1, including negative.

**CLM_FREQ**: not normal - poisson type

**HOME_VAL**: quasi-normal - skewed distribution with heavy tails

**HOMEKIDS*: Beta distribution

**INCOME**:  quasi-normal - skewed distribution with heavy tails

**KIDSDRIV**: Negative binomial / Poisson

**MVR_PTS**: Beta distribution

**TARGET_AMT**: Gamma distribution

**TIF**: Poisson distribution

**TRAVTIME**: quasi-normal - skewed distribution with heavy tails

**YOJ**: normal distribution with  heavy tail

**OLDCLAIM**: Poisson distribution

**CLM_FREQ**: Beta distribution

Some of the variables present a lot of zeros which could be explained as lack of data, and as such should be excluded, for example in HOMEVAL, while in others they are a rightful part of the distribution, and should be considered in the analysis, such as in INCOME, CLM_FREQ, HOMEKIDS, KIDSDRV, MVR_PTS,  OLDCLAIM, etc.

For the categorical features, we will displayed their distribution using bar charts.

```{r message=FALSE, warning=FALSE, echo=FALSE}
Train_Data2<-Train_Data %>% select_if(negate(is.numeric))
Train_Data2 %>%
  gather() %>%                            
  ggplot(aes(value)) +                    
    facet_wrap(~ key, scales = "free", nrow = 4) +  
   geom_bar() + coord_flip()
```

Some of the features have several sub-categories, like **CAR_TYPE**, **EDUCATION**, and **JOB**, while the other features are binary in nature. Interaction between these sub-categories and the continous variables to be taken into consideration while building models.

## Correlation matrix

Considering the number of variables and sub-categories within the discrete features, the correlation matrix visualization is challenging. We will then show two matrices one with numeric only and other with discrete variable. Analysis are based on the whole dataset, though.

We also ran 'pairs' a function that produces a matrix of scatterplots - not displayed here due to size.

```{r, echo=FALSE}
# par(mfcol = 1:2)
model.matrix(~0+., Train_Data1) %>%
  cor(use="pairwise.complete.obs") %>%
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
```

```{r, echo=FALSE, fig.height = 28, fig.width = 20}
model.matrix(~0+., Train_Data2) %>%
  cor(use="pairwise.complete.obs") %>%
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
```

Some observations from the above charts:

- positive correlation:    
Income and HomeVal    
Income and BlueBook    
SexF and CarType SUV      
Phd Degree and Job as Doctor
Master's Degree and Job as Lawyer
Income and Education
Income and Urbanicity Urban
     
- Negative correlation:      
Age and HomeKids    
HomeKids and CarAge  
Urbanicity Rural and Claim frequency
Urbanicity Rural and BlueBook    

## Evaluation dataset              
Procedures described above were also applied to the evaluation set.


# Data Prep

Looking at the plots we see we have to make a few changes to some variables. We'll make HOMEKIDS boolean instead of a factor. For the rows where AGE and CAR_AGE are less than zero, we make them equal to 0. For blank JOBS we label those as "Unknown". Finally, change Education to 1 if PhD and Masters.

```{r Homekids, echo=FALSE}
Train_Data$HOMEKIDS[Train_Data$HOMEKIDS != 0 ] <- 1
Eval_Data$HOMEKIDS[Eval_Data$HOMEKIDS != 0 ] <- 1
```

```{r Age, echo=FALSE}
Train_Data$CAR_AGE[Train_Data$AGE < 0 ] <- 0
Eval_Data$CAR_AGE[Eval_Data$AGE < 0 ] <- 0
```

```{r CarAge, echo=FALSE}
Train_Data$CAR_AGE[Train_Data$CAR_AGE < 0 ] <- 0
Eval_Data$CAR_AGE[Eval_Data$CAR_AGE < 0 ] <- 0
```

```{r JOB, echo=FALSE}
Train_Data$JOB <- as.character(Train_Data$JOB)
Train_Data$JOB[Train_Data$JOB == ""] <- "Unknown"
Train_Data$JOB <- as.factor(Train_Data$JOB)
Eval_Data$JOB <- as.character(Eval_Data$JOB)
Eval_Data$JOB[Eval_Data$JOB == ""] <- "Unknown"
Eval_Data$JOB <- as.factor(Eval_Data$JOB)
```

```{r Education, echo=FALSE}
Train_Data$EDUCATION <- ifelse(Train_Data$EDUCATION %in% c("PhD", "Masters"), 0, 1)
Eval_Data$EDUCATION <- ifelse(Eval_Data$EDUCATION %in% c("PhD", "Masters"), 0, 1)
```


## Missing Data

We have missing data for income, yoj, home_val, and car_age variables.

```{r}
summary(Train_Data)
```

We assume the missing data are Missing at Random and choose to impute. The reason we want to impute the missing data rather than replacing with mean or median because of large number of missing values. If we're replacing with mean or median on the large number of missing values, can result in loss of variation in data. We're imputing the missing data using the MICE package. The method of predictive mean matching (PMM) is selected for continuous variables.

```{r, echo=FALSE}
impute_train_data <- mice(Train_Data, m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_train_data) 
complete_train_data <- complete(impute_train_data,2)
impute_eval_data <- mice(Eval_Data, m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_eval_data) 
complete_eval_data <- complete(impute_eval_data,2)
```

# Building Models

## Binary Model 1 - All Variables

Our first model will seek to create a baseline using binary response variable, using a logistic regression model that contains all of our features. 

```{r, echo=FALSE}
m1b <- glm(factor(TARGET_FLAG) ~ ., family=binomial, data=subset(complete_train_data, select=-c(TARGET_AMT)))
summary(m1b)
```
The AIC result from the binomial model can be derived using the logit link function. 

## Binary Model 2 - Hand Pick Model


We can see that from saturated model above that variables *AGE*, *YOJ*, *PARENT1*, *SEX*, *EDUCATION*, *JOB*, and *RED_CAR* have p values greater than 0.05. These variables will be dropped to build the next model.

Also, we see some predictors are skewed and so we take log of them to build model 2.

```{r, echo=FALSE}
m2b <- glm(factor(TARGET_FLAG) ~ KIDSDRIV+HOMEKIDS+log(INCOME+1)+log(HOME_VAL+1)+MSTATUS+log(TRAVTIME)+CAR_USE+log(BLUEBOOK)+TIF	+CAR_TYPE+log(OLDCLAIM+1)+CLM_FREQ+REVOKED+MVR_PTS +log(CAR_AGE+1)+URBANICITY, family=binomial, data=subset(complete_train_data, select=-c(TARGET_AMT)))
summary(m2b)
```

## Binary Model 3 - Forward Step Model


We will now build a model using forward selection in order to compare if using forward selection is better than hand picking values to create a model.

We can use the same stepAIC function to build the third model. The forward selection approach starts from the null model and adds a variable that improves the model the most, one at a time, until the stopping criterion is met. We can see the result is different compared to the backward selection approach. The AIC is a little higher.

```{r, echo=FALSE}
m3b <- stepAIC(m1b, direction='forward', trace=FALSE)
summary(m3b)
```

## Binary Model 4 - Stepwise Step Model

We also can use the same stepAIC function to build the fourth model using stepwise regression. The stepwise regression method involves adding or removing potential explanatory variables in succession and testing for statistical significance after each iteration. This is exactly same result as the backward step model. 

```{r, echo=FALSE}
m4b <- stepAIC(m1b, direction='both', trace=FALSE)
summary(m4b)
```

Judging by AIC, the stepwise approach reduces the dimensionality and improves fit, given its lower estimated prediction error. This suggests that, in addition to being a simple model, the stepwise method works better to create an overall better fit to the data.

The analysis of deviance table shows further confirms that dropping these statistical insignificant variables  {*AGE*, *SEX*, *EDUCATION*, *RED_CAR*} in model 4.

```{r, echo=FALSE}
anova(m4b,m1b, test="Chi")
```

We next ran some multivariate regression models, using the features to predict the numeric response variable, TARGET_AMT, which gives the costs associated with a car's accident, creating a multiple linear regression model to predict the response variable.


## Multi Linear Regression Model 1 - All Variables

For the multi linear regression we want to know what is going to be the insurance cost if a person has crashed their car. We are going to build a multi linear regression model which includes all the data, from there we will keep the variables that have significance and use that to build subsequence models. We will first need to create a dataset specifically for a multi linear regression as we only care about if a customer has crashed their car. 

```{r, echo=FALSE}
mlr_data = complete_train_data %>%
  filter(TARGET_FLAG == 1) %>%
  dplyr::select(-1)
```

```{r, echo=FALSE}
mlr1 = lm(TARGET_AMT ~ ., data = mlr_data)
summary(mlr1)
```

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(mlr1)
```


Looking at the plot above we can see that the Residual vs Fitted graph has a large variance for a couple of outliers while a majority of the points have very low residuals. Also looking at the Normal Q-Q graph we can see that it is not a normal distribution. This is quite good for a model which utilizes all of the variables and we would like to see if we can improve the model by selecting variables that are significant.

Some variables that we would like to use for the next model are **KIDSDRIV**, **SEX**, **CAR_USE**, **REVOKED**, and **CAR_AGE**. These variables makes a lot of are usually thought of as the variables which can increase the cost of insurance

## Multi Linear Regression Model 2 - Hand Picking Variables

```{r, echo=FALSE}
mlr2 = lm(TARGET_AMT ~ KIDSDRIV + SEX + CAR_USE + REVOKED + CAR_AGE, data=mlr_data)
summary(mlr2)
```

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(mlr2)
```

With the second model we can see that we have a lower R Squared value in our new model compared to the first model which included all the variables. We can also see that there is still a large variance with the Residual vs Fitted plot. We will next try to use a stepwise function to find the best model from all the variables.

## Multi Linear Regression Model 3 - Stepwise Function

```{r, echo=FALSE}
mlr3 = stepAIC(mlr1, direction='backward', trace=FALSE)
summary(mlr3)
```

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(mlr3)
```

Using the backward step function we can see that the model choose **MSTATUS**, **SEX**, **BLUEBOOK**, **REVOKED**, and **MVR_PTS**. Looking at the residual vs fitted plot we still see a variance caused by outliers.

# Model Selection

## Logistic Regression

we will compare various metrics for all 4 models. We check models’ confusion matrix, accuracy, classification error rate, precision, sensitivity, specificity, F1 score, AUC, and ROC curves.

First, let's plot the ROC curves for all 4 models and then calculate the various metrics.

```{r echo=FALSE, message=FALSE}
par(mfrow=c(2,2))
plot(roc(complete_train_data$TARGET_FLAG,  predict(m1b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 1')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m2b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 2')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m3b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 3')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m4b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 4')
```

```{r echo=FALSE, message=FALSE, results=FALSE}
# comparing all models using various measures
CM1 <- confusionMatrix(as.factor(as.integer(fitted(m1b) > .5)), as.factor(m1b$y), positive = "1")
CM2 <- confusionMatrix(as.factor(as.integer(fitted(m2b) > .5)), as.factor(m2b$y), positive = "1")
CM3 <- confusionMatrix(as.factor(as.integer(fitted(m3b) > .5)), as.factor(m3b$y), positive = "1")
CM4 <- confusionMatrix(as.factor(as.integer(fitted(m4b) > .5)), as.factor(m4b$y), positive = "1")
```

```{r echo=FALSE, message=FALSE, results=FALSE}
Roc1 <- roc(complete_train_data$TARGET_FLAG,  predict(m1b, complete_train_data, interval = "prediction"))
Roc2 <- roc(complete_train_data$TARGET_FLAG,  predict(m2b, complete_train_data, interval = "prediction"))
Roc3 <- roc(complete_train_data$TARGET_FLAG,  predict(m3b, complete_train_data, interval = "prediction"))
Roc4 <- roc(complete_train_data$TARGET_FLAG,  predict(m4b, complete_train_data, interval = "prediction"))
```

```{r echo=FALSE, message=FALSE, results=FALSE}
metrics1 <- c(CM1$overall[1], "Class. Error Rate" = 1 - as.numeric(CM1$overall[1]), CM1$byClass[c(1, 2, 5, 7)], AUC = Roc1$auc)
metrics2 <- c(CM2$overall[1], "Class. Error Rate" = 1 - as.numeric(CM2$overall[1]), CM2$byClass[c(1, 2, 5, 7)], AUC = Roc2$auc)
metrics3 <- c(CM3$overall[1], "Class. Error Rate" = 1 - as.numeric(CM3$overall[1]), CM3$byClass[c(1, 2, 5, 7)], AUC = Roc3$auc)
metrics4 <- c(CM4$overall[1], "Class. Error Rate" = 1 - as.numeric(CM4$overall[1]), CM4$byClass[c(1, 2, 5, 7)], AUC = Roc4$auc)
```

```{r echo=FALSE}
kable(cbind(metrics1, metrics2, metrics3, metrics4), col.names = c("Model 1", "Model 2", "Model 3", "Model 4"))  %>% 
  kable_styling(full_width = T)
```

By looking at the ROC curves, model 1, 3, and 4 are showing the same area under curve value. So, it's hard to justify which model is the best. Fortunately, we have the various calculated metrics to provide us more details which model is the best. Based on that, we can say that the model 4 performs the highest in all metrics except Class. Error Rate. 

## Multi Linear Regression

We will be looking at all the models created and looking at the metrics like R Squared Value, RMSE, F-Statistics, and Residual Plots in order to determine which is the best model which represents our data. We will then compare the best model selected against the evaluation data set in order to see if the model truly represents the dataset

```{r, echo=FALSE}
m1.summary = summary(mlr1)
m2.summary = summary(mlr2)
m3.summary = summary(mlr3)
```

```{r, echo=FALSE}
m1.square = m1.summary$r.squared
m2.square = m2.summary$r.squared
m3.square = m3.summary$r.squared
```

```{r f-statistics, echo=FALSE}
m1.fstat = as.numeric(m1.summary$fstatistic[1])
m2.fstat = as.numeric(m2.summary$fstatistic[1])
m3.fstat = as.numeric(m3.summary$fstatistic[1])
```

```{r rsquare, echo=FALSE}
m1.r = m1.summary$adj.r.squared
m2.r = m2.summary$adj.r.squared
m3.r = m3.summary$adj.r.squared
```

```{r, echo=FALSE}
metrics1 = c('R Square'=m1.square, 'F Stat'=m1.fstat, 'R Adj Square'=m1.r)
metrics2 = c('R Square'=m2.square, 'F Stat'=m2.fstat, 'R Square'=m2.r)
metrics3 = c('R Square'=m3.square, 'F Stat'=m3.fstat, 'R Square'=m3.r)
```

```{r echo=FALSE}
kable(cbind(metrics1, metrics2, metrics3), col.names = c("Model 1", "Model 2", "Model 3"))  %>% 
  kable_styling(full_width = T)
```

Based on the table above the the best model would be model 3 based on the summary and the residual vs fitted plot

## Predictions

## Predictions

Logistics Model:
```{r}
prediction_binary = predict(m4b, complete_eval_data, type="response")
complete_eval_data$TARGET_FLAG = prediction_binary
complete_eval_data$TARGET_FLAG <- ifelse(complete_eval_data$TARGET_FLAG > 0.5, 1, 0)
print(head(complete_eval_data,10))
```

Multi Linear Model:
```{r}
prediction_linear = predict(mlr3, complete_eval_data)
complete_eval_data$TARGET_AMT = ifelse(complete_eval_data$TARGET_FLAG ==1, prediction_linear, 0)
print(head(complete_eval_data,10))
write.csv(complete_eval_data, 'predictions.csv')
```

# Appendix

```{r echo=TRUE, eval=FALSE}
Train_Data <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW4/insurance_training_data.csv")
Eval_Data <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW4/insurance-evaluation-data.csv")
Train_Data <- Train_Data[,-1]
str(Train_Data)
summary(Train_Data)
Train_Data$INCOME <- gsub(",","",(Train_Data$INCOME))
Train_Data$INCOME <- sub('.', '', Train_Data$INCOME)
Train_Data$INCOME <-trimws(Train_Data$INCOME, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$INCOME <- as.numeric(Train_Data$INCOME)
Train_Data$HOME_VAL <- gsub(",","",(Train_Data$HOME_VAL))
Train_Data$HOME_VAL <- sub('.', '', Train_Data$HOME_VAL)
Train_Data$HOME_VAL <-trimws(Train_Data$HOME_VAL, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$HOME_VAL <- as.numeric(Train_Data$HOME_VAL)
# 
Train_Data$BLUEBOOK <- gsub(",","",(Train_Data$BLUEBOOK))
Train_Data$BLUEBOOK <- sub('.', '', Train_Data$BLUEBOOK)
Train_Data$BLUEBOOK <-trimws(Train_Data$BLUEBOOK, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$BLUEBOOK <- as.numeric(Train_Data$BLUEBOOK)
# 
Train_Data$OLDCLAIM <- gsub(",","",(Train_Data$OLDCLAIM))
Train_Data$OLDCLAIM <- sub('.', '', Train_Data$OLDCLAIM)
Train_Data$OLDCLAIM <-trimws(Train_Data$OLDCLAIM, which = c("both"), whitespace = "[ \t\r\n]")
Train_Data$OLDCLAIM <- as.numeric(Train_Data$OLDCLAIM)
# 
summary(Train_Data[,c(7,9,16,20)])
colSums(is.na(Train_Data)) 
Train_Data1<-dplyr::select_if(Train_Data, is.numeric)
Train_Data1 %>%
  keep(is.numeric) %>%                     
  tidyr::gather() %>%                            
  ggplot(aes(value)) +                     
    facet_wrap(~ key, scales = "free") +  
    geom_density()
Train_Data$HOMEKIDS[Train_Data$HOMEKIDS != 0 ] <- 1
Eval_Data$HOMEKIDS[Eval_Data$HOMEKIDS != 0 ] <- 1
Train_Data$CAR_AGE[Train_Data$AGE < 0 ] <- 0
Eval_Data$CAR_AGE[Eval_Data$AGE < 0 ] <- 0
Train_Data$CAR_AGE[Train_Data$CAR_AGE < 0 ] <- 0
Eval_Data$CAR_AGE[Eval_Data$CAR_AGE < 0 ] <- 0
Train_Data$JOB <- as.character(Train_Data$JOB)
Train_Data$JOB[Train_Data$JOB == ""] <- "Unknown"
Train_Data$JOB <- as.factor(Train_Data$JOB)
Eval_Data$JOB <- as.character(Eval_Data$JOB)
Eval_Data$JOB[Eval_Data$JOB == ""] <- "Unknown"
Eval_Data$JOB <- as.factor(Eval_Data$JOB)
summary(Train_Data)
impute_train_data <- mice(Train_Data, m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_train_data) 
complete_train_data <- complete(impute_train_data,2)
impute_eval_data <- mice(Eval_Data, m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_eval_data) 
complete_eval_data <- complete(impute_eval_data,2)
m1b <- glm(factor(TARGET_FLAG) ~ ., family=binomial, data=subset(complete_train_data, select=-c(TARGET_AMT)))
summary(m1b)
m2b <- glm(factor(TARGET_FLAG) ~ KIDSDRIV+HOMEKIDS+log(INCOME+1)+log(HOME_VAL+1)+MSTATUS+log(TRAVTIME)+CAR_USE+log(BLUEBOOK)+TIF	+CAR_TYPE+log(OLDCLAIM+1)+CLM_FREQ+REVOKED+MVR_PTS +log(CAR_AGE+1)+URBANICITY, family=binomial, data=subset(complete_train_data, select=-c(TARGET_AMT)))
summary(m2b)
m3b <- stepAIC(m1b, direction='forward', trace=FALSE)
summary(m3b)
m4b <- stepAIC(m1b, direction='both', trace=FALSE)
summary(m4b)
anova(m4b,m1b, test="Chi")
mlr_data = complete_train_data %>%
  filter(TARGET_FLAG == 1) %>%
  dplyr::select(-1)
mlr1 = lm(TARGET_AMT ~ ., data = mlr_data)
summary(mlr1)
par(mfrow=c(2,2))
plot(mlr1)
mlr2 = lm(TARGET_AMT ~ KIDSDRIV + SEX + CAR_USE + REVOKED + CAR_AGE, data=mlr_data)
summary(mlr2)
par(mfrow=c(2,2))
plot(mlr2)
mlr3 = stepAIC(mlr1, direction='backward', trace=FALSE)
summary(mlr3)
par(mfrow=c(2,2))
plot(mlr3)
par(mfrow=c(2,2))
plot(roc(complete_train_data$TARGET_FLAG,  predict(m1b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 1')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m2b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 2')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m3b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 3')
plot(roc(complete_train_data$TARGET_FLAG,  predict(m4b, complete_train_data, interval = "prediction")), print.auc = TRUE, main='ROC Curve Model 4')
CM1 <- confusionMatrix(as.factor(as.integer(fitted(m1b) > .5)), as.factor(m1b$y), positive = "1")
CM2 <- confusionMatrix(as.factor(as.integer(fitted(m2b) > .5)), as.factor(m2b$y), positive = "1")
CM3 <- confusionMatrix(as.factor(as.integer(fitted(m3b) > .5)), as.factor(m3b$y), positive = "1")
CM4 <- confusionMatrix(as.factor(as.integer(fitted(m4b) > .5)), as.factor(m4b$y), positive = "1")
Roc1 <- roc(complete_train_data$TARGET_FLAG,  predict(m1b, complete_train_data, interval = "prediction"))
Roc2 <- roc(complete_train_data$TARGET_FLAG,  predict(m2b, complete_train_data, interval = "prediction"))
Roc3 <- roc(complete_train_data$TARGET_FLAG,  predict(m3b, complete_train_data, interval = "prediction"))
Roc4 <- roc(complete_train_data$TARGET_FLAG,  predict(m4b, complete_train_data, interval = "prediction"))
metrics1 <- c(CM1$overall[1], "Class. Error Rate" = 1 - as.numeric(CM1$overall[1]), CM1$byClass[c(1, 2, 5, 7)], AUC = Roc1$auc)
metrics2 <- c(CM2$overall[1], "Class. Error Rate" = 1 - as.numeric(CM2$overall[1]), CM2$byClass[c(1, 2, 5, 7)], AUC = Roc2$auc)
metrics3 <- c(CM3$overall[1], "Class. Error Rate" = 1 - as.numeric(CM3$overall[1]), CM3$byClass[c(1, 2, 5, 7)], AUC = Roc3$auc)
metrics4 <- c(CM4$overall[1], "Class. Error Rate" = 1 - as.numeric(CM4$overall[1]), CM4$byClass[c(1, 2, 5, 7)], AUC = Roc4$auc)
kable(cbind(metrics1, metrics2, metrics3, metrics4), col.names = c("Model 1", "Model 2", "Model 3", "Model 4"))  %>% 
  kable_styling(full_width = T)
m1.summary = summary(mlr1)
m2.summary = summary(mlr2)
m3.summary = summary(mlr3)
m1.square = m1.summary$r.squared
m2.square = m2.summary$r.squared
m3.square = m3.summary$r.squared
m1.fstat = as.numeric(m1.summary$fstatistic[1])
m2.fstat = as.numeric(m2.summary$fstatistic[1])
m3.fstat = as.numeric(m3.summary$fstatistic[1])
m1.r = m1.summary$adj.r.squared
m2.r = m2.summary$adj.r.squared
m3.r = m3.summary$adj.r.squared
metrics1 = c('R Square'=m1.square, 'F Stat'=m1.fstat, 'R Adj Square'=m1.r)
metrics2 = c('R Square'=m2.square, 'F Stat'=m2.fstat, 'R Square'=m2.r)
metrics3 = c('R Square'=m3.square, 'F Stat'=m3.fstat, 'R Square'=m3.r)
kable(cbind(metrics1, metrics2, metrics3), col.names = c("Model 1", "Model 2", "Model 3"))  %>% 
  kable_styling(full_width = T)
prediction_binary = predict(m4b, complete_eval_data, type="response")
complete_eval_data$TARGET_FLAG = prediction_binary
complete_eval_data$TARGET_FLAG <- ifelse(complete_eval_data$TARGET_FLAG > 0.5, 1, 0)
print(head(complete_eval_data,10))
prediction_linear = predict(mlr3, complete_eval_data)
complete_eval_data$TARGET_AMT = ifelse(complete_eval_data$TARGET_FLAG ==1, prediction_linear, 0)
print(head(complete_eval_data,10))
write.csv(complete_eval_data, 'predictions.csv')
```
