---
title: "Data 621 Homework 1"
output: pdf
---


```{r}
library(tidyverse)
```

# Data Preperation
```{r importing data set}
moneyball_training_data = read.csv("moneyball-training-data.csv")
```


```{r finding all NA}
na_count = sapply(moneyball_training_data, function(y) sum(is.na(y)))
na_count = data.frame(na_count)
na_count %>%
  arrange(desc(na_count)) %>%
  mutate(total_rows = nrow(moneyball_training_data)) %>%
  mutate(percent_missing = na_count / total_rows)
```
We can see that **TEAM_BATTING_HBP** is missing 91% of its data and **TEAM_BASERUN_CS** is missing around 34% of its data. This is a lot of data missing which is why those columns will be removing these. Based on online reading there is no definite cut of for how much data one should be missing before removing a column, but it is always better to have more data. With the other columns that are missing 10% of its data maybe we can fill those in with the mean and median, or entirely remove the row as there should be enough data left to train the model.

```{r removing column}
moneyball_subset = subset(moneyball_training_data, select=-c(TEAM_BATTING_HBP, TEAM_BASERUN_CS))
```

## Transformation
Checking all of the columns to see if they need any type of transformation on the column in order to create a linear line


