---
title: "DATA 621 Homework 5"
author: "Critical Thinking Group 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4

---
\pagebreak
\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Prof. Dr. Nasrin Khansari\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
DATA 621 – Business Analytics and Data Mining\\
\medskip
Home Work 5\\
\medskip
Prepared by:\\
\medskip
Critical Thinking Group  1\\ 
\medskip
Vic Chan\\ 
\smallskip
Gehad Gad\\
\smallskip
Evan McLaughlin\\  
\smallskip
Bruno de Melo\\
\smallskip
Anjal Hussan\\
\smallskip
Zhouxin Shi\\
\smallskip
Sie Siong Wong\\
\end{center}

\pagebreak

\newpage

# Introduction
```{r, message=F, warning=FALSE, echo=FALSE}
if (!require('pacman')) (install.packages('pacman'))
pacman::p_load('ggplot2', 'reporttools', 'dplyr', 'MASS', 'dplyr', 'psych', 'DataExplorer', 'mice', 'pscl', 'pander', 'tinytex')


theme_update(plot.title = element_text(hjust = 0.5), 
             axis.text.x = element_text(angle = 90, hjust = 1))
train <- read.csv('https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW5/wine-training-data.csv', header=T)
evaluation_data <- read.csv('https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW5/wine-evaluation-data.csv', header=T)
```

## Problem

Our goal is to explore, analyze and model a dataset containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine.  These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant.

A large wine manufacturer is studying the data in order to predict  the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.

The objective is to build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine.

\newpage

## 1. DATA EXPLORATION

Below we'll display a few basic EDA techniques to gain insight into our wine dataset.

### Basic Statistics

The data is 1.3 Mb in size. There are 12,795 rows and 15 columns (features). Of all 15 columns, 0 are discrete, 15 are continuous, and 0 are all missing. There are 8,200 missing values out of 191,925 data points.

```{r, echo=FALSE, warning=FALSE}
summary <- describe(train[,c(1:15)])[,c(2:5,8,9,11,12)]
#knitr::kable(summary)
print(summary)
```

### Histogram of Variables

```{r, echo=FALSE, warning=FALSE}
plot_histogram(train)
```

### Relationship of Predictors to Target

```{r, echo=FALSE, warning=FALSE}
plot_scatterplot(train[2:15,], "TARGET")
```

