---
title: "DATA 621 Homework 5"
author: "Critical Thinking Group 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '2'
    smart: no
  word_document:
    toc: yes
    toc_depth: '2'
---
\pagebreak
\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Prof. Dr. Nasrin Khansari\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
DATA 621 – Business Analytics and Data Mining\\
\medskip
Home Work 5\\
\medskip
Prepared by:\\
\medskip
Critical Thinking Group  1\\ 
\medskip
Vic Chan\\ 
\smallskip
Gehad Gad\\
\smallskip
Evan McLaughlin\\  
\smallskip
Bruno de Melo\\
\smallskip
Anjal Hussan\\
\smallskip
Zhouxin Shi\\
\smallskip
Sie Siong Wong\\
\end{center}

\pagebreak


```{r message=FALSE, warning=FALSE, echo=FALSE}
if (!require('ggplot2')) (install.packages('ggplot2'))
if(!require('dplyr')) (install.packages('dplyr'))
if(!require('tidyverse')) (install.packages('tidyverse'))
if(!require('purrr')) (install.packages('purrr'))
if(!require('mice')) (install.packages('mice'))
if(!require('DataExplorer')) (install.packages('DataExplorer'))
if(!require('MASS')) (install.packages('MASS'))
if(!require('caret')) (install.packages('caret'))
if(!require('stats')) (install.packages('stats'))
if(!require('pROC')) (install.packages('pROC'))
if(!require('kableExtra')) (install.packages('kableExtra'))
if(!require('gridExtra')) (install.packages('gridExtra'))
if(!require('ggcorrplot')) (install.packages("ggcorrplot"))
if(!require('fitdistrplus')) (install.packages("fitdistrplus"))
if (!require('pacman')) (install.packages('pacman'))
pacman::p_load('ggplot2', 'reporttools', 'dplyr', 'MASS', 'dplyr', 'psych', 'DataExplorer', 'mice', 'pscl', 'pander', 'tinytex')
```


\newpage

# Introduction

```{r, message=F, warning=FALSE, echo=FALSE}
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.text.x = element_text(angle = 90, hjust = 1))

train_data <- read.csv('https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW5/wine-training-data.csv', header=T)
eval_data <- read.csv('https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/HW5/wine-evaluation-data.csv', header=T)
```

## Problem

Our goal is to explore, analyze and model a dataset containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine.  These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant.

A large wine manufacturer is studying the data in order to predict  the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.

The objective is to build a count regression model to predict the number of cases of wine that will be sold given certain properties of the wine.

\newpage

# Data Exploration  

Below we'll display a few basic EDA techniques to gain insight into our wine dataset.

## Basic Statistics

The data is 1.3 Mb in size. There are 12,795 rows and 15 columns (features). Of all 15 columns, 0 are discrete, 15 are continuous, and 0 are all missing. There are 8,200 missing values out of 191,925 data points.

```{r, echo=FALSE, warning=FALSE}
summary <- describe(train_data[,c(1:15)])[,c(2:5,8,9,11,12)]
#knitr::kable(summary)
print(summary)
```

## Histogram of Variables

```{r, echo=FALSE, warning=FALSE}
plot_histogram(train_data)
```

## Relationship of Predictors to Target

```{r, echo=FALSE, warning=FALSE}
plot_scatterplot(train_data[2:15,], "TARGET")
```


# Data Preparation

## Identify Missing Values

We can see that the same variables for both train data set and evaluation data set contains missing values. The variable that contains the most missing values is the **STAR** followed by **Sulphates**, **Alcohol**, **ResidualSugar**, **TotalSulfurDioxide**, **FreeSulfurDioxide**, **Chlorides**, and **pH**.

````{r, echo=FALSE, warning=FALSE}
# Missing values in train data set
colSums(is.na(train_data[-c(1)])) %>% data.frame() %>% `colnames<-`("Missing Count") %>% tibble::rownames_to_column("Train Data Set Variales") 
plot_missing(train_data[-c(1)])

# Missing values in evaluation data set
colSums(is.na(eval_data[-c(1)])) %>% data.frame() %>% `colnames<-`("Missing Count") %>% tibble::rownames_to_column("Train Data Set Variales")
plot_missing(eval_data[-c(1,2)])
```

## Impute Missing Values

We assume the missing data are Missing at Random and choose to impute. The reason we want to impute the missing data rather than replacing with mean or median because of large number of missing values. If we're replacing with mean or median on the large number of missing values, can result in loss of variation in data. We're imputing the missing data using the MICE package. The method of predictive mean matching (PMM) is selected for continuous variables.

```{r, echo=FALSE, warning=FALSE}
# Impute both train and evaluation data set
impute_train_data <- mice(train_data[-c(1)], m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_train_data) 
impute_eval_data <- mice(eval_data[-c(1)], m=5, maxit=20, method='pmm', seed=321, print = FALSE)
densityplot(impute_eval_data)
```

Next, we take average values of the 5 imputed data set as a final train data set used for building models. 

```{r, echo=FALSE, warning=FALSE}
# Combine all 5 imputed data set and calculate the average value
complete_train_data <- (complete(impute_train_data,1) + complete(impute_train_data,2) + complete(impute_train_data,3) + complete(impute_train_data,4) + complete(impute_train_data,5))/5

complete_eval_data <- (complete(impute_eval_data,1) + complete(impute_eval_data,2) + complete(impute_eval_data,3) + complete(impute_eval_data,4) + complete(impute_eval_data,5))/5

# Plot box plot for each variable
complete_train_data %>% gather(variable, value) %>% 
  ggplot(aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=variable)) + facet_wrap( ~ variable, scales="free") + 
  theme(legend.position = "none", axis.text.y=element_blank()) + coord_flip()

# Plot density plot for each variable
complete_train_data %>% plot_density()

```

Because we'll use the Poisson or Negative Binomial regression to build count regression models in GLM approach, transformation for each variable to make them looks normal is not required. Diagnostics of actual outliers or influential points can be identified in the build models section through plots such as residuals vs fitted, standardized residuals vs fitted, etc.


# Build Models


# Model Selection


# Appendix

